---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

First let's read the data (It is assumed that the data files "activity.csv" is in the working directory)
```{r}
activity<-read.csv("activity.csv",header=TRUE,stringsAsFactor=FALSE)
```

Next steps is to convert the date column into POSIX date format for future processing
```{r}
activity$date<-as.POSIXct(activity$date)
```

## What is mean total number of steps taken per day?

For this analysis we will use the dplyr package. We compute the total steps per day with dplyr package
```{r,message=FALSE}
# Load dplyr library
library(dplyr)
```
```{r}
# Group by date and compute total
activity.g<-group_by(activity,factor(date))
tot.steps<-summarize(activity.g, sum(steps,na.rm=TRUE))
names(tot.steps)<-c("date","tot.steps")
head(tot.steps)
```
Histogram of total number of steps per day
```{r}
hist(tot.steps$tot.steps,xlab="Number of steps",main="Total number of steps per day")
```
Mean of total number of step taken per day
```{r}
mean(tot.steps$tot.steps,na.rm=TRUE)
```

Median of total number of steps taken per day
```{r}
median(tot.steps$tot.steps,na.rm=TRUE)
```

## What is the average daily activity pattern?
Using dplyr we group by intervals then compute the average per interval across all date
```{r}
activity.g<-group_by(activity,factor(interval))
aver.steps<-summarize(activity.g,mean(steps,na.rm=TRUE))
names(aver.steps)<-c("interval","aver.steps")
```
We can now plot the time series of average steps per interval

```{r}
plot(as.vector(aver.steps$interval),aver.steps$aver.steps,xlab="5-min intervals",
     ylab="Average number of steps",type="l",col="blue")
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
