---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

First let's read the data (It is assumed that the data files "activity.csv" is in the working directory)
```{r}
activity<-read.csv("activity.csv",header=TRUE,stringsAsFactor=FALSE)
```

Next steps is to convert the date column into POSIX date format for future processing
```{r}
activity$date<-as.POSIXct(activity$date)
```

## What is mean total number of steps taken per day?

For this analysis we will use the dplyr package. We compute the total steps per day with dplyr package
```{r,message=FALSE}
# Load dplyr library
library(dplyr)
```
```{r}
# Group by date and compute total
activity.g<-group_by(activity,factor(date))
tot.steps<-summarize(activity.g, sum(steps,na.rm=TRUE))
names(tot.steps)<-c("date","tot.steps")
head(tot.steps)
```
Histogram of total number of steps per day
```{r}
hist(tot.steps$tot.steps,xlab="Number of steps",main="Total number of steps per day")
```

Mean of total number of step taken per day
```{r}
mean(tot.steps$tot.steps,na.rm=TRUE)
```

Median of total number of steps taken per day
```{r}
median(tot.steps$tot.steps,na.rm=TRUE)
```

## What is the average daily activity pattern?
Using dplyr we group by intervals then compute the average per interval across all date
```{r}
activity.g<-group_by(activity,factor(interval))
aver.steps<-summarize(activity.g,mean(steps,na.rm=TRUE))
names(aver.steps)<-c("interval","aver.steps")
```

We can now plot the time series of average steps per interval

```{r}
plot(as.vector(aver.steps$interval),aver.steps$aver.steps,xlab="5-min intervals",
     ylab="Average number of steps",type="l",col="blue")
```

To extract the 5-min interval with the maximum number of steps we use the following code
```{r}
# Get the indice of the maximum number of steps
ind.max<-which.max(aver.steps$aver.steps)
# Extract the 5-min interval with the maximum number of steps
int.max<-as.vector(aver.steps$interval)[ind.max]
# Print to screen
print(as.numeric(int.max))
```

## Imputing missing values
We will first calculate and report the total number of NA (rows with NAs)
```{r}
# Extract indices of row with NA values
ind.na<-which(is.na(activity$steps)==TRUE)

# count how many indices there is
nb.na<-length(ind.na)

# Print the number of rows with NA
print(nb.na)
```
The NA will be replaced by the average number of steps per 5-min interval. First we will extract the NA rows and the rows without NA in two seperate dataframes
```{r}
# extract all the rows with NA in a new data.frame
activ.na<-activity[ind.na,]

# Extract all the rows without NA in a new data.frame
activ<-activity[-ind.na,]

```
Next we will merge the previously computed average number of steps per 5-min interval with the NA dataframe using the "interval" columns as a common data field

```{r}
merged.na<-merge(activ.na,aver.steps,by.x="interval",by.y="interval")

# Rearrange the merged result to prepare it to be appended to the dataframe without NA
merged.na<-merged.na[,c(4,3,1)]

# Rename the first column
names(merged.na)[1]<-"steps"

# Merge dataframes to create the new dataframe with imputted NAs
activity.new<-rbind(activ,merged.na)
head(activity.new)
```

We recompute the total number of steps per day using the new dataframe with the imputed missing data:
```{r}
# Group by date and compute total
activity.g<-group_by(activity.new,factor(date))
tot.steps<-summarize(activity.g, sum(steps,na.rm=TRUE))
names(tot.steps)<-c("date","tot.steps")
head(tot.steps)
```
Histogram of total number of steps per day after missing values imputting
```{r}
hist(tot.steps$tot.steps,xlab="Number of steps",main="Total number of steps per day")
```

Mean of total number of step taken per day after missing values imputting
```{r}
mean(tot.steps$tot.steps,na.rm=TRUE)
```

Median of total number of steps taken per day after missing values imputting
```{r}
median(tot.steps$tot.steps,na.rm=TRUE)
```
The new histogram is less skewed (closer to a normal distribution) as evidenced by the mean and the median being equal. Also the mean is higher (9350 before missing values imputting vs 1076 afterwards).

## Are there differences in activity patterns between weekdays and weekends?
We will first convert dates to days and create a code for weekdays and weekends: 1 for weekdays and 2 for weekends.
```{r}

# Converting dates to days
date.days<-weekdays(activity.new$date)

# Creating indices for weekend
ind.wknd<-which(date.days=="Saturday" | date.days=="Sunday")

# Adding new factor
activity.new$fact.d<-1
activity.new$fact.d[ind.wknd]<-2

```


